
## Load required R packages

```{r packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(arsenal)
library(kableExtra)
library(nlme)
library(emmeans)
library(splines)
library(clubSandwich)
library(A4LEARN)

formatp <- function(x) case_when(
  x < 0.001 ~ "p<0.001",
  x > 0.01 ~ Hmisc::format.pval(x, digits=2, eps=0.01, nsmall=2),
  TRUE ~ Hmisc::format.pval(x, digits=3, eps=0.001, nsmall=3))
```

```{r settings, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, 
  warning = FALSE, cache = TRUE)

a4Palette <- c("#0072b5", "#bc3c29")
scale_colour_discrete <-
  function(...) scale_colour_manual(..., values = a4Palette)
scale_fill_discrete <-
  function(...) scale_fill_manual(..., values = a4Palette)

theme_set(theme_bw() +
  theme(legend.background = element_rect(fill="transparent")))
```

## Data manipulations

```{r data-manipulation}
# Outcomes collected at Visit 1
V1OUTCOME <- A4LEARN::ADQS %>%
  dplyr::filter(VISITCD == "001") %>%
  select(BID, QSTESTCD, QSSTRESN) %>%
  pivot_wider(values_from = QSSTRESN, names_from = QSTESTCD)

# Outcomes collected at Visit 6
V6OUTCOME <- A4LEARN::ADQS %>%
  dplyr::filter(VISITCD == "006") %>%
  select(BID, QSTESTCD, QSSTRESN) %>%
  pivot_wider(values_from = QSSTRESN, names_from = QSTESTCD)

SUBJINFO <- A4LEARN::SUBJINFO %>%
  left_join(V6OUTCOME, by = "BID") %>%
  left_join(V1OUTCOME %>%
      select(BID, CDRSB, CFITOTAL, CFISP, CFIPT, ADLPQPT, ADLPQSP), 
    by = "BID")

# Filter A4LEARN::ADQS for PACC collected in the blinded phases among mITT population
ADQS_PACC <- A4LEARN::ADQS %>%
  dplyr::filter(MITTFL== 1) %>%
  dplyr::filter(EPOCH == "BLINDED TREATMENT" | AVISIT == "006") %>%
  dplyr::filter(QSTESTCD == "PACC") %>%
  rename(PACC = QSSTRESN) %>%
  select(BID, ASEQNCS, TX, ADURW, TX, AGEYR, 
    AAPOEGNPRSNFLG, EDCCNTU, SUVRCER, QSVERSION, PACC) %>%
  mutate(TX = factor(TX, levels = c("Placebo", "Solanezumab"))) %>%
  na.omit()
```

## Baseline characteristics of the A4 trial

```{r table1, results = "asis"}
A4labels <- list(TX = "Treatment group", AGEYR = "Age (y)", 
  EDCCNTU = "Education (y)", SUVRCER = "FBP SUVr", AMYLCENT = "FBP Centiloid", 
  LMIIa = "LM Delayed Recall",  MMSE = "MMSE", 
  CFITOTAL = "CFI Combined", ADLPQSP = "ADL Partner", CDRSB = "CDR-SB",
  SEX = "Sex", RACE = "Racial categories", ETHNIC = "Ethnicity", 
  MARITAL = "Marital Status",  WRKRET = "Retirement Status", 
  APOEGNPRSNFLG = "APOE e4", APOEGN = "APOE Genotype")

table1 <- tableby(TX ~ AGEYR + EDCCNTU + SEX + RACE + ETHNIC + MARITAL + WRKRET + 
    SUVRCER + AMYLCENT + chisq(APOEGN) + chisq(APOEGNPRSNFLG) + 
    PACC + LMIIa + MMSE + CFITOTAL + ADLPQSP + CDRSB, 
  data = SUBJINFO %>% dplyr::filter(MITTFL== 1), 
  control = tableby.control(test=TRUE,
    stats.labels = list(Nmiss = "Missing")))

summary(table1, labelTranslations = A4labels, digits = 1,
  title = "Characteristics of the A4 participants by randomized treatment group (Modified Intention-to-Treat Population of those who received at least one dose of solanezumab or placebo and underwent assessment for the primary end point).")
```

## Primary analyis of the PACC

Preclinical Alzheimer Cognitive Composite (PACC, @donohue2014preclinical) data was modeled using natural cubic splines as described in @donohue2023natural. The fixed effects included the following terms: (i) spline basis expansion terms (two terms), (ii) spline basis expansion terms-by-treatment interaction (two terms), (iii) PACC test version administered, (iv) baseline age, (v) education, (vi) APOE4 Carrier Status (yes/no), and (vii) baseline florbetapir cortical SUVr. The model is constrained to not allow a difference
between treatment group means at baseline. The following variance-covariance structures were assumed in sequence until a model converged: (i) heterogeneous unstructured, (ii) heterogeneous Toeplitz, (iii) heterogeneous autoregressive order 1, (iv) heterogeneous compound symmetry, and (v) compound symmetry. The first structure, heterogeneous unstructured, did not converge. Here we fit the second, heterogeneous Toeplitz, which did converge.

```{r pacc-spaghetti-plot, fig.cap="Spaghetti plot of PACC data over time by treatment group."}
ggplot(ADQS_PACC, aes(x=ADURW, y=PACC, color=TX)) +
  geom_line(aes(group = BID), alpha=0.2) +
  theme(legend.position = "inside", legend.position.inside = c(0.2, 0.2)) +
  xlab("Weeks since Randomization") +
  scale_x_continuous(breaks = seq(0, max(ADQS_PACC$ADURW), by = 24))
```

```{r pacc-spline-model-fit}
# Spline basis expansion functions
# CAUTION: these function depend on the ADQS_PACC data frame object in the 
# global environment. If ADQS_PACC changes, so do these funtions
ns21 <- function(t){
  as.numeric(predict(splines::ns(ADQS_PACC$ADURW, df=2,
    Boundary.knots = c(0, max(ADQS_PACC$ADURW))), t)[,1])
}
ns22 <- function(t){
  as.numeric(predict(splines::ns(ADQS_PACC$ADURW, df=2,
    Boundary.knots = c(0, max(ADQS_PACC$ADURW))), t)[,2])
}

# GLS model fit:
pacc_fit <- gls(PACC ~ 
    I(ns21(ADURW)) + I(ns22(ADURW)) +
    (I(ns21(ADURW)) + I(ns22(ADURW))):TX + 
    AGEYR + AAPOEGNPRSNFLG + EDCCNTU + SUVRCER + QSVERSION,
  data = ADQS_PACC,
  weights = varIdent(form = ~ 1 | ASEQNCS),
  correlation = corARMA(form = ~ ASEQNCS | BID, p = 10))
```

```{r pacc-summarize-spline-model-1}
ref_grid(pacc_fit,
  at = list(ADURW = c(0,240), TX = levels(ADQS_PACC$TX)),
  vcov. = clubSandwich::vcovCR(pacc_fit, type = "CR2") %>% as.matrix(), 
  data = ADQS_PACC, 
  mode = "satterthwaite") %>%
  emmeans(~ ADURW | TX) %>%
  pairs(reverse = TRUE) %>%
  as_tibble() %>%
  mutate(p.value = formatp(p.value)) %>%
  kable(caption = "Mean PACC change from baseline at week 240 by treatment 
    group estimated from spline model.", digits = 2, booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

```{r pacc-summarize-spline-model-2}
contrast240 <- ref_grid(pacc_fit, 
  at = list(ADURW = 240, TX = levels(ADQS_PACC$TX)),
  vcov. = clubSandwich::vcovCR(pacc_fit, type = "CR2") %>% as.matrix(), 
  data = ADQS_PACC, 
  mode = "satterthwaite") %>%
  emmeans(specs = "TX", by = "ADURW") %>%
  pairs(reverse = TRUE, adjust = "none")

contrast240 %>% 
  as_tibble() %>%
  left_join(contrast240 %>% 
      confint() %>%
      as_tibble() %>%
      select(contrast, lower.CL, upper.CL), by="contrast") %>%
  relocate(p.value, .after = last_col()) %>%
  mutate(p.value = formatp(p.value)) %>%
  kable(caption = "Mean PACC group change from baseline at week 240 by treatment 
    group estimated from spline model.", digits = 2, booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

```{r pacc-mean-plot, fig.cap="Mean PACC per treatment group over time as estimated by spline model. Shaded region depicts 95% confidence intervals."}
ref_grid(pacc_fit, 
  at = list(ADURW = seq(0, 312, by=12), TX = levels(ADQS_PACC$TX)),
  vcov. = clubSandwich::vcovCR(pacc_fit, type = "CR2") %>% as.matrix(), 
  data = ADQS_PACC, 
  mode = "satterthwaite") %>%
  emmeans(specs = "TX", by = "ADURW") %>%
  as_tibble() %>%
  ggplot(aes(x=ADURW, y=emmean)) +
  geom_line(aes(color=TX)) +
  geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL, fill = TX), alpha = 0.2) +
  scale_x_continuous(breaks = seq(0, 312, by=24)) +
  ylab("Mean PACC with 95% confidence intervals") +
  xlab("Weeks since Randomization") +
  theme(legend.position = 'inside', legend.position.inside = c(0.2, 0.2))
```

## References
